---
title: "Reproducible Research: Peer Assessment 1"
author: Hans U Deck
date: "9 6 2021"
output: 
  html_document:
   keep_md: true

---


#### **Loading and preprocessing the data**
```{r Loading and preprocessing data}
library(ggplot2)
data <- read.csv("repdata_data_activity/activity.csv")
head(data)
str(data)
summary(data)

sum <- aggregate(data$steps, by = list(date = data$date), FUN = sum, ia.rm = TRUE)
sum$date <- as.Date(sum$date)
head(sum)
str(sum)
summary(sum)
  
```

The dataframe (sum) includes the the total steps each day.

#### **What is mean total number of steps taken per day?**
```{r steps per day}
steps <- sum$x
h <- hist(
        steps,
        main = "Histogram  Steps per Day",
        breaks = 60,
        xlim = c(0, 25000),
        xlab = "Total Steps / Day"
)

text(
        h$mids,
        h$counts,
        labels = h$counts,
        adj = c(0.5,-0.5),
        cex = 0.7
)

```



Calculation the **mean** value:

```{r}
mean <- round(mean(sum$x, na.rm = TRUE), digits = 0)
mean
```

**The mean value is `r format(mean, digits=8)`**
`

Calculation the **median** value:

```{r}
median <- (median(sum$x, na.rm = TRUE))
median
```

**The mean value is `r format(median, digits=8)`**



#### **What is the average daily activity pattern?**

```{r}
mean <- aggregate(data$steps, by = list(date = data$date), FUN = mean, na.rm = TRUE)
mean$x <- round(mean$x, digits = 2)
mean$date <- as.Date(mean$date)
p <- ggplot(mean, (aes(date, x))) +
        geom_line() +
        labs(x = "Date", y = " Average Steps per day", title = "Time Serines average number of steps per day") +
        theme_light()

p

```



#### **Imputing missing values**

```{r}
CountNA <- sum(is.na(data$steps))
CountNA
```

The Total numbers of rows with NA is  `r CountNA`

```{r}
data1 <- data
data1$steps[is.na(data1$steps)] <- mean(data1$steps, na.rm = TRUE)
head(data1)
summary(data1)
```

Filling the missing values with mean function

```{r}

sum1 <- aggregate(data1$steps, by = list(date = data1$date), FUN = sum, ia.rm = TRUE)
sum1$date <- as.Date(sum1$date)
head(sum1)
str(sum1)
summary(sum1)

steps <- sum1$x
h <- hist(
        steps,
        main = "Histogram  Steps per Day",
        breaks = 60,
        xlim = c(0, 25000),
        xlab = "Total Steps / Day"
)

text(
        h$mids,
        h$counts,
        labels = h$counts,
        adj = c(0.5,-0.5),
        cex = 0.7
)
```

```{r}
meanValue <- round(mean(sum1$x), digits = 0)
meanValue
medianValue <- round(median(sum1$x), digits = 0)
medianValue
```





`**The mean Value is `r format(meanValue, digits = 8) ` and the median Value is `r format(medianValue, digit = 8)` **





#### **Are there differences in activity patterns between weekdays and weekends?**

```{r}
library(timeDate)

sumtest <- sum1
sumtest$days <- isWeekday(sumtest$date)

weekday <- subset(sumtest, sumtest$days == TRUE)
weekday$days <- "Weekday"

weekend <- subset(sumtest, sumtest$days == FALSE)
weekend$days <- "Weekend"

week <- rbind(weekday, weekend)

week <- week[order(week$date),]
week$date <- as.Date(week$date)

p <- ggplot(week, aes(date, x, color = days)) +
        geom_line() +
        labs(x = "Date", y = " Average Steps per day", title = "Time Serines average number of steps per day") +
        facet_wrap( ~ days, ncol = 1) +
        theme_light()
p


```

